Postural changes mediate children's visual access to social information

### 2018-01-14 writeup-cogsci-copy/cogsci2018.Rmd
got rid of merge conflict inserts

### Detector evaluation
In order to evaluate the performance of these detectors, we constructed a "gold set" of frames

>>>>

### Detector evaluation
In order to evaluate the performance of these detectors, we constructed a "gold set" of frames
Evaluating face detectors (text here about how gold standard was selected)

\begin{table}[t]
  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
  \begin{center}
    \begin{tabular}{l|ccc|ccc}
      \hline
       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
      % \null Model & P & R & F & P & R & F  \\
      \hline
      % MTCNN & .84  & 1 &  .91 & .95 & .75 & .85   \\
      % OpenPose & .99 & .87 & .93 & .71 & .92 & .8 \\
    \hline
    \end{tabular}
  \end{center}
\end{table}

Evaluating hand/wrist detectors (text here about how gold standard was selected)

### 2018-01-14 writeup-cogsci-copy/cogsci2018.tex
got rid of merge conflict inserts

\begin{abstract}

@@

Children are remarkably skilled language learners, connecting arbitrary
labels (``cup'') with specific visual concepts at a rapid pace through
the first two years of life. However, children do not learn words in a
vacuum but rather in a rich social environment, where social cues
provided by speakers (e.g., eye-gaze) provide strong scaffolding for
this learning process. Further, children's ability to effectively
process these social cues may be a key factor in their early language
development. In a longitudinal study, children's level of joint
engagement with their mother was found to predict both their receptive
and productive vocabularies (Carpenter, Nagell, \& Tomasello (1998)).
More recently, 10 month-olds who follow an adult's gaze in an
experimental context have larger vocabularies at 18 months (Brooks \&
Meltzoff (2005)) and throughout the second year of life
(({\textbf{???}})).

However, as children are learning their first words, their view of the
world is also changing radically (Adolph \& Berger, 2007). Infant's
motor abilities improve dramatically near the end of the first year of
life, allowing them to locomote independently and to determine what they
see. During spontaneous play in a laboratory playroom, toddlers are more
likely to look at the floor while crawling than while walking (Franchak
et al., 2011), when they have full visual access to their environment
and the people in it (Kretch, Franchak, and Adolph, 2014).

@@

To obtain this final sample, we tested 95, excluding 59 children for the
following reasons: 20 for technical issues related to the headcam, 15
for failing to wear the headcam, 10 for fewer than 4 minutes of headcam
footage, 5 for having multiple adults present, 5 for missing CDI data, 2
for missing scene camera footage, 1 for fussiness, and one excluded for
sample symmetry. All inclusion decisions were made independent of the
results of subsequent analyses.

@@

Measuring the availability of caregiver's faces across development was
an important goal of the study. We used a total of three face detection
systems to explore this hypothesis and to avoid the cost of
hand-annotating every frame. Results for the evaluation of these three
systems are shown in Table X.

@@

The second algorithm was based on the work by (CITEXX) using multi-task
cascaded convolutional neural networks (MTCNNs). The system was built
using a novel cascaded CNN-based framework for joint detection and
alignment, built to perform well in real-world environments where
varying illuminations and occlusions are present. We used a Tensorflow
implementation of this algorithm provided by
(\url{https://github.com/davidsandberg/facenet})

The third algorithm was a CNN-based pose detector that provided the
locations of 18 body parts (ears, nose, wrists, etc) (CITEXX). The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose. For the purposes of
our investigation we only made use of the body parts relevant to the
face (ears, eyes, nose).

@@

\begin{table}[t]
  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
  \begin{center}
    \begin{tabular}{l|ccc|ccc}
      \hline
       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
      % \null Model & P & R & F & P & R & F  \\
      \hline
      % Viola-Jones & .55  & .38 &  .45 & .89 & .74 & .81   \\
      % MTCNN & .86 & .78 & .81 & .93 & .76 & .83 \\
      % OpenPose-Faces & .86 & .78 & .81 & .93 & .76 & .83 \\
      % OpenPose-Wrists & .86 & .78 & .81 & .93 & .76 & .83 \\
      % OpenPose-Hands & .86 & .78 & .81 & .93 & .76 & .83 \\
    \hline
    \end{tabular}
  \end{center}
\end{table}

@@

\section{Changes in Posture and
Orientation}\label{changes-in-posture-and-orientation}

{\centering \includegraphics{figs/unnamed-chunk-4-1}

@@

\caption[R plot]{R plot}\label{fig:unnamed-chunk-4}
\end{figure}
\end{CodeChunk}

\section{Changes in Access to Faces}\label{changes-in-access-to-faces}

@@

{\centering \includegraphics{figs/unnamed-chunk-5-1}

@@

\caption[Proportion face (left) and hand (right) detections as a function of participant's age]{Proportion face (left) and hand (right) detections as a function of participant's age. Note that the axes are different; hands were detected overall much less frequently in this dataset.}\label{fig:unnamed-chunk-5}
\end{figure*}
\end{CodeChunk}

@@

\caption[Proportion of face detections as a function of infant's posture, using the MTCNN3 detections (left panel) and OpenPose detections (right panel)]{Proportion of face detections as a function of infant's posture, using the MTCNN3 detections (left panel) and OpenPose detections (right panel)}\label{fig:faceByPostureFig}
\end{figure*}
\end{CodeChunk}

{\centering \includegraphics{figs/faceByOrientFig-1}

\caption[Proportion of face detections as a function of caregiver's orientation,using the MTCNN3 detections (left panel) and OpenPose detections (right panel)]{Proportion of face detections as a function of caregiver's orientation,using the MTCNN3 detections (left panel) and OpenPose detections (right panel)}\label{fig:faceByOrientFig}
\end{figure*}
\end{CodeChunk}

@@

\section{Changes in Access to Hands}\label{changes-in-access-to-hands}

Next, we examined whether we would see similar trends in infant and
children's access to hands.

>>>>

\begin{abstract}
The ability to process social cues (e.g., eye gaze) is a critical
component of children's early language and cognitive development.
However, as children reach their first birthday, they begin to locomote
themselves, exploring their visual environment in an entirely new way.
How do these postural changes affect children's access to the social
information relevant for word-learning? Here, explore this question by
using head-mounted cameras to record infants' (8-16 months of age)
egocentric visual perspective and by using computer-vision algorithms to
detect the proportion of faces and hands in infants' environments. We
find that infant's postural changes largely mediate infants' accesses to
both faces and hands, suggesting that these postural and locomotive
developments facilitate infants' emerging linguistic and social
capacities.

@@

Children are remarkably skilled language learners, connecting arbitrary
labels (``cup'') with specific visual concepts at a rapid pace through
the first two years of life. However, children do not learn words in a
vacuum but in a rich social environment, where social cues provided by
speakers (e.g., eye-gaze) provide strong scaffolding for this learning
process. Indeed, children's ability to effectively process these social
cues may be a key factor in their early language development. For
example,in a longitudinal study, children's level of joint engagement
with their mother was found to predict both their receptive and
productive vocabularies (Carpenter, Nagell, \& Tomasello (1998)). More
recently, 10 month-olds who follow an adult's gaze in an experimental
context have larger vocabularies at 18 months (Brooks \& Meltzoff
(2005)) and throughout the second year of life (({\textbf{???}})).

However, as children are learning their first words, their view of the
world is also changing radically (Adolph \& Berger, 2007). Infant's
motor abilities improve dramatically near the end of the first year of
life, allowing them to locomote independently and to determine what they
see. These motor changes have drastic consequences for what children
see: crawling and walking infants have radically different views of the
world. During spontaneous play in a laboratory playroom, toddlers are
more likely to look at the floor while crawling than while walking
(Franchak et al., 2011); walking vs.~crawling infants tend to have full
visual access to their environment and the people in it (Kretch,
Franchak, and Adolph, 2014).

@@

To obtain this final sample, we tested 95 children, excluding 59
children for the following reasons: 20 for technical issues related to
the headcam, 15 for failing to wear the headcam, 10 for fewer than 4
minutes of headcam footage, 5 for having multiple adults present, 5 for
missing CDI data, 2 for missing scene camera footage, 1 for fussiness,
and one excluded for sample symmetry. All inclusion decisions were made
independent of the results of subsequent analyses.

@@

Measuring the availability of caregiver's faces across development was
an important goal of the study. We thus used a total of three face
detection systems to explore this hypothesis and to avoid the cost of
hand-annotating every frame. We first measure the performance of the
most commonly-used and widely avlaialbe face detection algorithmsm
(Viola-Jones). We use this as a benchmark for performance, as while it
can achieve impressive accuracy in some situations, it is is notoriously
bad at dealing with occluded faces (({\textbf{???}})). We next
capitalized on recent improvements in computer vision, testing the
performance of two other algorithms that extract face information in
distinct ways. Rsults for the evaluation of these three systems are
shown in Table XX.

The second of these algorithms (OpenPose) also captures information
about the ``skeleton'' of people in a given frame, outputting
information about the entire set of body parts that were detected. This
allowed us to extract hand information as well as wrist information. We
analyze wrist detections as a good proxy for the presence of a hand. We
do so first because we did not capture the entire visual field
experienced by the infants/children. Thus evne though the video may only
show the presence of a wrist, the caregiver's hand may be within the
child's field of view. Second, we do so because hands are often occluded
by objects when caregivers are interacting with children, yet still
visually accessible by the child and part of their joint interaction.
For example,if a caregiver was holding a toy and presenting it to the
infant, their wrists may be visible but their hand may not necessarily
be (as it is occluded by the toy).

@@

The second algorithm was based on the work by (CITEXX) using multi-task
cascaded convolutional neural networks (MTCNNs). The system was built
using a novel cascaded CNN-based framework for joint detection and
alignment, built to perform well in real-world environments where
varying illuminations and occlusions are present. We used a Tensorflow
implementation of this algorithm provided by
(\url{https://github.com/davidsandberg/facenet}). Like Viola-Jones, this
detector provided information about the presence of a face as well as
it's size and position.

The third algorithm was a CNN-based pose detector that provided the
locations of 18 body parts (ears, nose, wrists, etc) (CITEXX). The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose. For the purposes of
our investigation we only made use of the body parts relevant to the
face (ears, eyes, nose). We operationalized face detections as any
frames containing a nose and either the left or right eye; we
operationalized hand detections as any frames containing any of the 21
points detected on either the left or the right hand, and we
operationalized wrist detections any frames containing either the left
or the right wrist.

@@

In order to evaluate the performance of these detectors, we constructed
a ``gold set'' of frames Evaluating face detectors (text here about how
gold standard was selected)

\begin{table}[t]
  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
  \begin{center}
    \begin{tabular}{l|ccc|ccc}
      \hline
       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
      % \null Model & P & R & F & P & R & F  \\
      \hline
      % MTCNN & .84  & 1 &  .91 & .95 & .75 & .85   \\
      % OpenPose & .99 & .87 & .93 & .71 & .92 & .8 \\
    \hline
    \end{tabular}
  \end{center}
\end{table}

Evaluating hand/wrist detectors (text here about how gold standard was
selected)

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
   \hline
\end{tabular}
\caption{TABLE PLACEHOLDER} 
\end{table}

@@

\subsection{Changes in Posture and
Orientation}\label{changes-in-posture-and-orientation}

\begin{CodeChunk}
\begin{figure}[h]

{\centering \includegraphics{figs/orientation-1}

@@

\caption[Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)]{Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel).}\label{fig:orientation}
\end{figure}
\end{CodeChunk}

We noted characteristic changes in infants posture and orientation
across this developmental time range. The proportion of time infant's
spent sitting decreased with age, and the proportion of time infant's
spent standing increased with age. Both 8-month-olds and 12-month-olds
psent equivalent amounts of time either lying/crawling, which was
markedly decreased in the 16-month-olds, who spent most of their time
either sitting or standing (see Figure \ref{fig:posture}). We also
observed characteristic changes in children's orientation relative to
their caregivers: the 8-month-olds spent more time with their caregiver
behind them supporting their sitting positions (see Figure
\ref{fig:orientation}).

\begin{CodeChunk}
\begin{figure}[h]

{\centering \includegraphics{figs/posture-1}

\caption[Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)]{Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel).}\label{fig:posture}
\end{figure}
\end{CodeChunk}

\section{Changes in Access to Faces \&
Hands}\label{changes-in-access-to-faces-hands}

We first examined the proportion of face and hand detections across age;
a full summary can be seen in Figure \ref{fig:detsbyAge}. We observed a
slight u-shaped function both when analyzing the output of the MTCNN and
OpenPose detectors, such that 12-month-olds may experience less faces
than 8 or 12-month-olds. Next, we observed that hand detections
increased with the age of the participant, though they were relatively
infrequent in this dataset. As expected, we found that wrist detections
we relatively more frequent and also appeared to increase somewhat with
age, suggesting that caregivers hands may be just outside of the field
of view of our camera or that they were often occluded by toys.

Next, we turned to our hypothesis of interest, examinig the impact of
postural and locomotive developments on children's visual access to
faces and hands. Overall, we ofund that children's posture was a major
factor both in how many faces they saw during the play session. Infants
who were sitting saw more faces than infants who were lying down or
being carried, while infants who were standing saw the most faces. We
observed a similar pattern with respect to hands: children who were
sitting or standing tended to see more hands/wrists than children who
were lying down or held.

We next examined how the child's orientation relative to their caregiver
impacted their visual access to faces and hands. Overlal, we saw that
children who were far away from their caregiver were more likely to see
faces and hands than children who were close to their caregiver; this
was true within all age groups and for all detectors.

@@

{\centering \includegraphics{figs/detByAge-1}

@@

\caption[Proportion face (left) and hand (right) detections as a function of participant's age]{Proportion face (left) and hand (right) detections as a function of participant's age. Note that the axes are different; hands were detected overall much less frequently in this dataset.}\label{fig:detByAge}
\end{figure*}
\end{CodeChunk}

@@

\caption[Proportion of face/hand detections as a function of infant's posture (top panels) and orientation (bottom panels), using the MTCNN3 face detections and OpenPose face detections (left two panels) and using OpenPose hand detections and the OpenPose wrist detections (right two panels)]{Proportion of face/hand detections as a function of infant's posture (top panels) and orientation (bottom panels), using the MTCNN3 face detections and OpenPose face detections (left two panels) and using OpenPose hand detections and the OpenPose wrist detections (right two panels).}\label{fig:faceByPostureFig}
\end{figure*}
\end{CodeChunk}

\subsection{Model}\label{model}

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -8.775 & 0.236 & -37.259 & 0.000 \\ 
  Age & 0.879 & 0.166 & 5.292 & 0.000 \\ 
  Posture-Prone  & 1.215 & 0.110 & 11.082 & 0.000 \\ 
  Posture-Sit & 1.943 & 0.108 & 18.000 & 0.000 \\ 
  Posture-Stand  & 2.175 & 0.107 & 20.260 & 0.000 \\ 
  Orient.-Close  & 0.608 & 0.059 & 10.371 & 0.000 \\ 
  Orient.-Far & 2.244 & 0.060 & 37.222 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.} 
\end{table}
